---
title: "SQL and beyond the infinite"
output: html_document
---

## Filtering Rows and Columns

You can filter in SQL columns by SELECT and rows by WHERE. Example from the Lahman baseball database, download it and sqldf package first. 

```{r message=FALSE, warning=FALSE}
library(Lahman)
library(sqldf)
```

Next, enter this code to see see hr totals for 1927 Yankees. Instead of a massive 100,000 entry list, we could write: 

```{r}
query<-"SELECT playerID,teamID,yearID,HR FROM Batting
WHERE teamID='NYA' and yearID=1927"
sqldf(query)
```

Going further, heres an example of Yankees with over 40 home runs.

```{r}
query<-"SELECT playerID,teamID,yearID,HR FROM Batting
WHERE teamID='NYA' and HR>40"
sqldf(query)
```
Next we have Yankees with more than 40 home runs AND less and 60 strikeouts. A much more difficult feat. 
```{R}
query1<-"SELECT playerID,teamID,yearID,HR FROM Batting
WHERE teamID='NYA' and HR>40 and SO<60"
sqldf(query1)
```
Finally we come to Phillies with more than 30 home runs throughout the entire 70's. 
```{R}
query2<-"SELECT playerID,teamID,yearID,HR FROM Batting
WHERE teamID='PHI' and HR>30 and yearID>1970 and yearID<1979"
sqldf(query2)
```

Multiple pieces of data can always be linked with 'and'. Also don't forget to HIGHLIGHT WHAT YOU'RE RUNNING.

## Order By

The function, ORDER BY, can greatly refine oure query's, let's take a look.

```{r}
query3<-"SELECT playerID,yearID,HR 
FROM Batting
WHERE HR>50
ORDER BY HR DESC"

sqldf(query3)
```


## Aggregation

Query's can be refined even further through aggregation. Sum's are added to the SELECT line and we introduce the GROUP BY function. 

```{r}
query4<-"SELECT playerID,yearID,avg(HR)FROM Batting
GROUP BY playerID
HAVING avg(HR)>30
ORDER BY avg(HR) DESC"

sqldf(query4)
```

## Joins 

Joins are called joins because you combine the data in two separate tables. This makes query's much more user freindly. For example here is a list of the full team names for where Babe Ruth played,

```{r}
query5<-"SELECT playerID,Batting.yearID,name,Batting.HR
FROM Batting INNER JOIN Teams
ON Batting.teamID=Teams.teamID and Batting.yearID=Teams.yearID
WHERE playerID='ruthba01'"

sqldf(query5)
```

Here's another example of the full names of players that hit over fifty homeruns.

```{r}
query6<-"SELECT nameFirst,nameLast,teamID,yearID,HR
FROM Batting INNER JOIN Master
ON Batting.playerID=Master.playerID
WHERE HR>50
ORDER BY HR"

sqldf(query6)