---
title: "SQL and beyond the infinite"
output: html_document
---

## Filtering Rows and Columns

You can filter in SQL columns by SELECT and rows by WHERE. Example from the Lahman baseball database, download it and sqldf package first. 

```{r message=FALSE, warning=FALSE}
library(Lahman)
library(sqldf)
```

Next, enter this code to see see hr totals for 1927 Yankees. Instead of a massive 100,000 entry list, we could write: 

```{r}
query<-"SELECT playerID,teamID,yearID,HR FROM Batting
WHERE teamID='NYA' and yearID=1927"
sqldf(query)
```

Going further, heres an example of Yankees with over 40 home runs.

```{r}
query<-"SELECT playerID,teamID,yearID,HR FROM Batting
WHERE teamID='NYA' and HR>40"
sqldf(query)
```
Next we have Yankees with more than 40 home runs AND less and 60 strikeouts. A much more difficult feat. 
```{R}
query<-"SELECT playerID,teamID,yearID,HR FROM Batting
WHERE teamID='NYA' and HR>40 and SO<60"
sqldf(query)
```
Finally we come to Phillies with more than 30 home runs throughout the entire 70's. 
```{R}
query<-"SELECT playerID,teamID,yearID,HR FROM Batting
WHERE teamID='PHI' and HR>30 and yearID>1970 and yearID<1979"
sqldf(query)
```

Multiple pieces of data can always be linked with 'and'. Also don't forget to HIGHLIGHT WHAT YOU'RE RUNNING.
